---
title: "Fold Change Differential Peaks"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/home/mina")
source('~/FoldChange_THOR/R/create_count_matrix.R')
source('~/FoldChange_THOR/R/calculate_fold_change.R')
source('~/FoldChange_THOR/R/calculate_pvalues.R')
source('~/FoldChange_THOR/R/plot_foldChange_pValues.R')
source('~/FoldChange_THOR/R/plot_foldChange_counts.R')
```

## 1) Create count matrix
### 1a.) create .saf file
For using featureCounts package, we should first create a .saf file, containing list of regions of peaks based on THOR-exp-<date>-diffpeaks.narrowPeak file.
run the following command on the terminal:

```{bash}
echo -e "GeneID\tChr\tStart\tEnd\tStrand" > TMM_THOR-diffpeaks.saf
awk -v FS='\t' -v OFS='\t' 'FNR > 1 { print $4, $1, $2, $3, $6 }' ~/FoldChange_THOR/data/TMM_THOR-diffpeaks.narrowPeak >> ~/FoldChange_THOR/data/TMM_THOR-diffpeaks.saf
```

### 1b.) Execute featureCounts
Based of data type which can be 'Single-End' or 'Paired-End', file path of .saf file, and path of .bam files:

```{r}
data_type = 'Single-End'
bam_files_file_path = '~/FoldChange_THOR/data/bwa/mergedLibrary'
saf_file_path = '~/FoldChange_THOR/data/TMM_THOR-diffpeaks.saf'
count_matrix = create_count_matrix(data_type, bam_files_file_path, saf_file_path)
pander::pander(head(count_matrix, c(6, 6)))
```
```{r}
colnames(count_matrix)
```

## 2) Calculate fold changes
Based on indices of two biological groups:

```{r}
grp1 = c(1:10)
grp2 = c(11:19)
fold_changes = calculate_fold_change(count_matrix, grp1, grp2)
head(fold_changes)
```

## 3) plot fold changes in p-values

You can also plot fold changes in p-values:

```{r}
raw_pvalues = calculate_pvalues(count_matrix, grp1, grp2)
plot_foldChange_pValues(count_matrix, fold_changes, raw_pvalues, pvalue_thr = 0.05, lfc_thr = 0.58)
```

## 4) plot fold changes in normalized counts

You can also plot fold changes in normalized counts:

```{r}
plot_foldChange_counts(fold_changes, count_matrix, lfc_thr = 0.58)
```

